####Transpose

1.Show that A$^{T}$A $\neq$ AA$^{T}$

For any given m$\times$n matrix A, A$^{T}$ is a n$\times$m matrix where each element a$_{ij}$ in A$^{T}$ is equal to a$_{ji}$ in matrix A.

A = $\begin{bmatrix}a_{11} & a_{12} & a_{13}\\a_{21} & a_{22} & a_{23}\\a_{31} & a_{32} & a_{33}\end{bmatrix}$
and A$^{T}$ = $\begin{bmatrix}a_{11} & a_{21} & a_{31}\\a_{12} & a_{22} & a_{32}\\a_{13} & a_{23} & a_{33}\end{bmatrix}$

$\because$ All elements in A$^{T}$ don't eaqual to corresponding elements in A except elements on the diagonal. 
$\therefore$ A$\neq$A$^{T}$

Based on Properties of matrix multiplication,
$\because$ AB$\neq$BA for any given matrix A, B where A$\neq$B
$\therefore$ A$^{T}$A $\neq$ AA$^{T}$<br/><br/>

2.For a special type of square matrix A, we get A$^{T}$A = AA$^{T}$. Under what conditions could this be true?  

To satisfy A$^{T}$A = AA$^{T}$, A must satisfy A = A$^{T}$ so that AA$^{T}$ = AA = A$^{T}$A$^{T}$ = A$^{T}$A
and therefore, for each a$_{ij}$ in A must equal to a$_{ji}$.

For example:  
A = $\begin{bmatrix}1 & 2 & 3\\2 & 1 & 4\\3 & 4 & 1\end{bmatrix}$
and A$^{T}$ = $\begin{bmatrix}1 & 2 & 3\\2 & 1 & 4\\3 & 4 & 1\end{bmatrix}$ = A

I = $\begin{bmatrix}1 & 0 & 0\\0 & 1 & 0\\0 & 0 & 1\end{bmatrix}$
and I$^{T}$ = $\begin{bmatrix}1 & 0 & 0\\0 & 1 & 0\\0 & 0 & 1\end{bmatrix}$ = I <br/><br/><br/><br/>


####LU Decomposition

Write an R function to factorize a square matrix A into LU
```{r}
LUDecomposition <- function(matrixA){
  r <- nrow(matrixA)
  matrixL <- diag(r)#create an identity matrix of the same size as A
  matrixU <- matrixA
  i <- 2
  j <- 1
  k <- 1
  while(k <= r) {
    while (i <= r) {
      if (matrixU[i,j]!=0) {
        multiplier <- (-matrixU[i,j]/matrixU[k,j])
        matrixL[i,j] <- -multiplier
            while (j <= r) {
              matrixU[i,j] <- matrixU[i,j] + matrixU[k,j]*multiplier
              j <- j+1
            }
        }
          i <- i+1
          j <- 1
      }
    k <- k+1
    i <- k+1
    j <- k
  }
  matrixList <- list(matrixL,matrixU)
  return (matrixList)
}
```

Test the function with a 3 x 3 matrix
```{r}
testA <- matrix(c(2,1,-6,4,-4,-9,-4,3,5),nrow = 3)
testL <- LUDecomposition(testA)[[1]]
testU <- LUDecomposition(testA)[[2]]
testA
testL
testU
testL%*%testU
```